---
title: "Assignment3"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 3
*Completed by *place your name here*
### Create a RMarkdown respository:
1. **- README (10 pt) file in top level directory.** 
+ The readme will contain your name*
2. **- An MIT License (10 pt)**
3. **--original**  Directory containing assignment named original
4. **original/index.Rmd**    R file containing the text here without editing
5. ** index.Rmd**  R markup file containing your completed homework


### Question 1

#### Question 1 Background
We will use example data from R, the iris dataset.  For the assignment, I'll ask for you to complete things from the diamonds or iris dataset. These are preloaded datasets

data(diamonds)
summary(diamonds)
  
```{r}
  library('rmarkdown')
  library("ggplot2")
  diamonds<-diamonds
  iris<-iris
  summary(iris)
```


##### Question 1 Background: Analysis of our iris dataset
```{r}
# Calculates the mean values for the aggregates given by the Species column in the iris data set.
 iris_mean <- aggregate(iris[,1:4], by=list(Species=iris$Species), FUN=mean)
# Calculates the standard deviations for the aggregates given by the Species column in the iris data set.
 iris_sd <- aggregate(iris[,1:4], by=list(Species=iris$Species), FUN=sd)
 convertDF <- function(df=df, mycolnames=c("Species", "Values", "Samples")) { myfactor <- rep(colnames(df)[-1], each=length(df[,1]));
# Defines function to convert data frames into ggplot2-friendly format.
mydata <- as.vector(as.matrix(df[,-1])); df <- data.frame(df[,1], mydata, myfactor); colnames(df) <- mycolnames; return(df) }
# Converts iris_mean.
 df_mean <- convertDF(iris_mean, mycolnames=c("Species", "Values", "Samples"))
# Converts iris_sd.
 df_sd <- convertDF(iris_sd, mycolnames=c("Species", "Values", "Samples"))
# Define standard deviation limits.
 limits <- aes(ymax = df_mean[,2] + df_sd[,2], ymin=df_mean[,2] - df_sd[,2])
```

#### **Question 1** (8 Points): 
1. (2pt) Please put the Mean Petal Length into mean_petal_length variable
+ *Provide your answer here*
2. (2pt) Please put the Max Petal Length max_petal_length variable
+ *Provide your answer here*
3. How many vectors (columns) are there in Iris and what are there names (4pt)?
+ *Provide your answer here*

### Question 2

#### Question 2 Background

##### Question 2 Background: Examples with *ggplot2*
  Ploting two vectors (columns) in form of a scatter plot against each other and places into a png image called *twovectors.png* in the images directory.  Make sure you have made your directory
```{r}
dir.create('images')
ggplot(iris, aes(Sepal.Length, Sepal.Width)) + geom_point()
png('images/twovectors.png')
ggplot(iris, aes(Sepal.Length, Sepal.Width)) + geom_point()
dev.off()
```

##### Question 2 Background Example 1: Plots larger dots and colors them with default color scheme.
```{r example1}
example1 <- ggplot(iris, aes(Sepal.Length, Sepal.Width)) + geom_point(aes(color = Species), size=4)
return(example1)
```

##### Question 2 Background Example 2: Colors dots with custom scheme.
```{r example2}
example2 <- ggplot(iris, aes(Sepal.Length, Sepal.Width)) + geom_point(aes(color = Species), size=4) + ylim(2,4) + xlim(4,8) + scale_color_manual(values=rainbow(10))
return(example2)
```

##### Question 2 Background Example 3: Adds a regression line to the plot.
```{r example3}
example3 <- ggplot(iris, aes(Sepal.Length, Sepal.Width)) + geom_point() + stat_smooth(method="lm", se=FALSE) + ylim(2,4) + xlim(4,8) + scale_color_manual(values=rainbow(10))
return(example3)
```
##### Question 2 Background Example 4: Plots three line plots, one for each sample in Species column.
```{r example4}
example4 <- ggplot(iris, aes(Sepal.Length, Sepal.Width)) + geom_line(aes(color=Species), size=1) + facet_wrap(~Species, ncol=1)
return(example4)
```

##### Question 2 Background Example 5: Plots with q plot
Qplot is another way to do ggplot2, but a bit simpler
```{r example5}
example5<-qplot(Sepal.Length, Petal.Length, data = iris, color = Species, size = Petal.Width)
return(example5)
```


##### Question 2 Background Example 6: same with ggplot
```{r example6}
example6<-ggplot(iris, aes(x = Sepal.Length, y = Petal.Length, color=Species)) + geom_point() + facet_wrap(~Species) + ggtitle("Plot of sepal length vs. petal length")
return(example6)
```

#### **Question 2** (20 Points): 
Using diamonds place graph of carat (x-axis) vs. price (y-axis), color by clarity, faceet by cut with a title



#### **Question 3** (20 Points): 

##### Question 3 Background Building a Widget
*Lets create our inputs (remember to make these code)*

inputPanel(
  selectInput("var1", label = "Choose Xaxis:", choices = c("Sepal.Length","Petal.Length"), selected = "Petal.Length"),
  selectInput("var2",  label = "Choose Yaxis:", choices = c("Sepal.Width","Petal.Width"), selected = "Petal.Width")
)

*Lets create our outputs (remember to make these code)*

renderPlot({
  x_axis <- input$var1
  y_axis <- input$var2
  species <- "Species"
   example13<-ggplot(iris, aes_string(x = x_axis, y = y_axis, color=species)) + geom_point() + facet_wrap(~Species) + ggtitle("Plot of sepal length vs. petal length")
   example13
}) 

```{r  echo = FALSE}
inputPanel(
  selectInput("var1", label = "Choose Xaxis:", choices = c("Sepal.Length","Petal.Length"), selected = "Petal.Length"),
  selectInput("var2",  label = "Choose Yaxis:", choices = c("Sepal.Width","Petal.Width"), selected = "Petal.Width")
)
```


```{r  echo = FALSE}
renderPlot({
  x_axis <- input$var1
  y_axis <- input$var2
  species <- "Species"
   example13<-ggplot(iris, aes_string(x = x_axis, y = y_axis, color=species)) + geom_point() + facet_wrap(~Species) + ggtitle("Plot of sepal length vs. petal length")
   example13
}) 
```
##### **Question 3** (20 Points): 
Make a similar shiny app with diamonds, allowing changing of axis.


